Notes for Week 2/23/19 - 2/29/19

### 2019-06-07 notes/weekly_notes/2.29.19.NotesForWeek.tex
weekly notes updated

\documentclass{article}
\usepackage{amsmath}
\usepackage{amsfonts, amsthm, amssymb}
\usepackage{graphicx}
\usepackage[colorlinks]{hyperref}
\usepackage[parfill]{parskip}
\usepackage{algpseudocode}
\usepackage{algorithm}
\usepackage{enumerate}
\usepackage[shortlabels]{enumitem}

@@

%%% Matrices
\newcommand{\Xbf}{\mathbf{X}}
\newcommand{\Ybf}{\mathbf{Y}}
\newcommand{\Zbf}{\mathbf{Z}}

@@

@@

Consider distributions $\Pbb$ and $\Qbb$ supported on $\D \subset \Reals^d$ which are absolutely continuous with density functions $f$ and $g$, respectively. For fixed $n \geq 0$, Let $\Zbf = (z_1, \ldots,z_n)$, where for $i = 1,\ldots,t$, $z_i \sim \frac{\Pbb + \Qbb}{2}$ are independent. Given $\Zbf$, for $i = 1,...,t$ let
\begin{equation*}
\ell_i = 
\begin{cases}
1~ \text{with probability $\frac{f(z_i)}{f(z_i) + g(z_i)}$} \\
-1~ \text{with probability $\frac{g(z_i)}{f(z_i) + g(z_i)}$}
\end{cases}
\end{equation*} 
be conditional independent labels, and write
\begin{equation*}
1_X = 
\begin{cases}
1,~ l_i = 1\\
0,~ \text{otherwise}
\end{cases}
1_Y = 
\begin{cases}
1,~ l_i = -1 \\
0,~ \text{otherwise.}
\end{cases}
\end{equation*}
We will write $\Xbf = \set{x_1, \ldots,x_{N_X}} := \set{z_i: \ell_i = 1}$ and similarly $\Ybf = \set{y_1, \ldots,y_{N_Y}} := \set{y_i: \ell_i = -1}$, where $N_X$ and $N_Y$ are of course random but $N_X + N_Y = n$.

@@

For $r \geq 0$, define the \emph{$r$-graph} $G_r = (V,E_r)$ to have vertex set $V = \set{1,\ldots,t}$ and edge set $E_r$ which contains the pair $(i,j)$ if and only if $\norm{z_i - z_j}_2 \leq r$. Let $D_{G_r}$ denote the incidence matrix of $G_r$.

Define the \emph{$r$-Laplacian Smooth} test statistic to be
\begin{equation*}
T_{LS} = \sup_{\theta: \norm{D_{G_r}\theta}_2 \leq C_{n,r}} \dotp{\theta}{\frac{1_X}{N_X} - \frac{1_Y}{N_Y}}
\end{equation*}

@@

\begin{lemma}[Spectral similarity of $r$-graph to grid]
	\label{lem: spectral_similarity_rgraph_grid}
	Fix $r \geq 2 \left(\frac{\log t}{n}\right)^{1/d} + (\frac{1}{n})^{1/d}$, and let $\ell(r,n) = \biggl(\sqrt{d} r n^{1/d} + 2\sqrt{d}(\log n)^{1/d}\biggr)^{1/2}$. For any $\theta \in \Reals^n$, the following relations hold:
	\begin{equation}
	\label{eqn: spectral_similarity_rgraph_grid}
	\frac{\norm{D_{G_r}\theta}_2}{\ell(t)} \leq \norm{D_{grid}\theta}_2 \leq \norm{D_{G_r}\theta}_2
	\end{equation}
	with probability at least $1 - n^{-\alpha}$ where $\alpha = c_1 (\log n)^{1/2}$ for some constant $c_1 > 0$.
\end{lemma}
\begin{proof}
	We begin by mapping the data $\Zbf$ to the grid points $[\kappa]^d$ in such a way that as little mass as possible is disturbed:
	\begin{lemma}
		\label{lem: mass_transport_mapping}
		There exists a bijective mapping $T: \Zbf \to [\kappa]^d$ for $\kappa = t^{1/d}$ such that
		\begin{equation*}
		\max_{i} \norm{T(z_i) - z_i}_2 \leq C\left(\frac{\log t}{t}\right)^{1/d}
		\end{equation*}
		with probability at least $1 - n^{-\alpha}$ where $\alpha = c_1 (\log n)^{1/2}$ for some constant $c_1 > 0$.
	\end{lemma}
	Hereafter, we assume there exists $T$ such that Lemma \ref{lem: mass_transport_mapping} holds.

	We first prove the second bound in \eqref{eqn: spectral_similarity_rgraph_grid}.  Consider grid points $k ~ k'$ connected in the grid graph. Then, there exist $z_i$ and $z_j$ such that $T(z_i) = k$ and $T(z_j) = k'$. By the triangle inequality,
	\begin{align*}
	\norm{z_i - z_j}_2 & \leq \norm{T(z_i) - z_i}_2 + \norm{T(z_i) - T(z_j)}_2 + \norm{T(z_j) - z_j}_2 \\
	& \leq 2C\left(\frac{\log t}{t}\right)^{1/d} + \frac{1}{t^{1/d}}
	\end{align*}
	and so by our choice of $r$, $i \sim j$ in $G_r$.

	Now, we turn to the first bound. Assume $i \sim j$ in the graph $G_r$. By a similar set of steps to the above, we have
	\begin{equation*}
	\norm{T(z_i) - T(z_j)}_2 \leq 2C\left(\frac{\log t}{t}\right)^{1/d} + r
	\end{equation*}
	As a result, using the simple relation $\norm{x}_1 \leq \sqrt{d} \norm{x}_2$ for any $x \in \Rd$, we have
	\begin{equation*}
	\norm{T(z_i) - T(z_j)}_1 \leq \sqrt{d}(2C\left(\frac{\log t}{t}\right)^{1/d} + r)
	\end{equation*}
	Since each edge in the grid graph is of length $n^{1/d}$, it is easy to see that there exists a path between $T(z_i)$ and $T(z_j)$ in $G_{grid}$, $P(T(Z_i) \to T(Z_j))$ with no more than
	\begin{equation*}
	\frac{\sqrt{d}(2C\left(\frac{\log t}{t}\right)^{1/d} + r)}{t^{1/d}}
	\end{equation*}
	edges. The bound follows by Lemma \ref{lem: graph_ordering}.
\end{proof}

\section{Poincare inequalities}

Let $G$ and $\widetilde{G}$ be undirected, unweighted graphs over vertex set $V$, with edge sets $E_G$ and $E_{\widetilde{G}}$, respectively. Let $\widetilde{\mathcal{P}}$ be the space of all paths over $E_{\widetilde{G}}$; that is, $\mathcal{P}$ consists of $\widetilde{P} \in \widetilde{\mathcal{P}}$
\begin{equation*}
P = (\widetilde{e}_1, \ldots, \widetilde{e}_m) \tag{$\widetilde{e_i} \in E_{\widetilde{G}}$}
\end{equation*}
for some integer $m \geq 1$. We say $\abs{P} = m$ is the \emph{path length}.

\begin{lemma}[Poincare inequality for general graphs.]
	Define a mapping $\gamma: E_G \to \mathcal{P}$ where for each $e = (\ell,\ell')$ in $E_G$
	\begin{equation*}
	\gamma(e) = ((\ell,u), \ldots, (v,\ell'))
	\end{equation*}
	meaning $e$ is mapped to a path which begins at $\ell$ and ends at $\ell'$. Then
	\begin{equation*}
	G \preceq \widetilde{G} \cdot \max_{e \in E_G} \abs{\gamma(e)}  \cdot b_{\gamma}
	\end{equation*}
	where $b_{\gamma}$ is a bottleneck parameter given by
	\begin{equation*}
	b_{\gamma} = \max_{\widetilde{e} \in E_{\widetilde{G}}} \abs{\set{e \in E: \widetilde{e} \in \widetilde{P}_e}}
	\end{equation*}
\end{lemma}
\begin{proof}
	Let $G_e = (V, \set{e})$ and $P_e = (V, \set{\widetilde{e}: \widetilde{e} \in \gamma(e)})$ be the graphs associated with $e$ and $\gamma(e)$, respectively. By Lemma \ref{lem: path_poincare}, we have
	\begin{equation*}
	G_{e} \preceq \abs{P_e} P_e
	\end{equation*}
	Summing over all $e \in E_G$, we obtain
	\begin{align*}
	G & \preceq \sum_{e \in E_G} \abs{P_e} P_e \\
	& \preceq \max_{e \in E_G} \abs{\gamma(e)} \sum_{e \in E_G} P_e \\
	& \preceq \max_{e \in E_G} \abs{\gamma(e)} b_{\gamma}\cdot \widetilde{G}
	\end{align*}
\end{proof}

@@

>>>>

\documentclass{article}
\usepackage{amsmath}
\usepackage{amsfonts, amsthm, amssymb}
\usepackage{graphicx}
\usepackage[colorlinks]{hyperref}
\usepackage[parfill]{parskip}
\usepackage{algpseudocode}
\usepackage{algorithm}
\usepackage{enumerate}
\usepackage[shortlabels]{enumitem}
\usepackage{fullpage}

@@

%%% Matrices
\newcommand{\X}{X} % no bold
\newcommand{\Y}{Y} % no bold
\newcommand{\Z}{Z} % no bold

@@

%%% Misc
\newcommand{\grid}{\mathrm{grid}}

@@

Consider distributions $\Pbb$ and $\Qbb$ supported on $\D \subset \Reals^d$ which are absolutely continuous with density functions $f$ and $g$, respectively. For fixed $n \geq 0$, let $\Z = (z_1, \ldots,z_n)$, where for $i = 1,\ldots,n$, $z_i \sim \frac{\Pbb + \Qbb}{2}$ are independent. Given $\Z$, for $i = 1,...,n$ let
\begin{equation*}
\ell_i = 
\begin{cases}
1~ \text{with probability $\frac{f(z_i)}{f(z_i) + g(z_i)}$} \\
-1~ \text{with probability $\frac{g(z_i)}{f(z_i) + g(z_i)}$}
\end{cases}
\end{equation*} 
be conditionally independent labels, and write
\begin{equation*}
1_X = 
\begin{cases}
1,~ l_i = 1\\
0,~ \text{otherwise}
\end{cases}
1_Y = 
\begin{cases}
1,~ l_i = -1 \\
0,~ \text{otherwise.}
\end{cases}
\end{equation*}
We will write $\X = \set{x_1, \ldots,x_{N_X}} := \set{z_i: \ell_i = 1}$ and similarly $\Y = \set{y_1, \ldots,y_{N_Y}} := \set{z_i: \ell_i = -1}$, where $N_X$ and $N_Y$ are of course random but $N_X + N_Y = n$.

@@

For $r \geq 0$, define the \emph{$r$-graph} $G_r = (V,E_r)$ to have vertex set $V = \set{1,\ldots,t}$ and edge set $E_r$ which contains the pair $(i,j)$ if and only if $\norm{z_i - z_j}_2 \leq r$. Let $D_{r}$ denote the incidence matrix of $G_r$.

For a critical radius $C_{n,r}$ to be determined later, define the \emph{$r$-Laplacian Smooth} test statistic to be
\begin{equation*}
T_{LS} = \sup_{\theta: \norm{D_{r}\theta}_2 \leq C_{n,r}} \dotp{\theta}{\frac{1_X}{N_X} - \frac{1_Y}{N_Y}}
\end{equation*}

@@

\begin{lemma}[Spectral similarity of $r$-graph to grid]
	\label{lem: spectral_similarity_rgraph_grid}
	Fix $r \geq 2 C\left(\frac{\log n}{n}\right)^{1/d} + \left(\frac{1}{n}\right)^{1/d}$, where $C > 0$ is a universal constant, and let 
	\begin{equation*}
	\sigma_{r,n} =  d^{d + 1/2} n^{2 + 1/d} \left(2C\left(\frac{\log n}{n}\right)^{1/d} + r \right)^{2d + 1}
	\end{equation*}

	For any $\theta \in \Reals^n$, there exists a permutation $\Pi: \Reals^d \to \Reals^d$ such that the following relations hold:
	\begin{equation}
	\label{eqn: spectral_similarity_rgraph_grid}
	\frac{\norm{D_{G_r}\theta}_2}{\sigma_{r,n}} \leq \norm{D_{grid} (\Pi \theta)}_2 \leq \norm{D_{G_r}\theta}_2
	\end{equation}
	with probability at least $1 - n^{-\alpha}$ where $\alpha = c_1 (\log n)^{1/2}$ for some constant $c_1 > 0$.
\end{lemma}

Lemma \ref{eqn: spectral_similarity_rgraph_grid} relies heavily on theory regarding optimal transportation matchings between two sets of discrete points, this case $\Z$ and $V_{grid}$.

\begin{lemma}
	\label{lem: mass_transport_mapping}
	There exists a bijection $T: \Z \to V_{\grid}$ such that
	\begin{equation*}
	\max_{i \in [n]} \norm{T(z_i) - z_i}_2 \leq C\left(\frac{\log n}{n}\right)^{1/d}
	\end{equation*}
	with probability at least $1 - n^{-\alpha}$, where $\alpha = c_1 (\log n)^{1/2}$ and $c_1, C > 0$ are universal constants.
\end{lemma}

The upper bound of \eqref{eqn: spectral_similarity_rgraph_grid} follows easily.
\begin{proof}[Upper bound of \eqref{eqn: spectral_similarity_rgraph_grid}]
	Assume there exists $T$ such that Lemma \ref{lem: mass_transport_mapping} holds.

	Let $k, k' \in [\kappa]^d$ satisfy $\frac{k}{\kappa} ~ \frac{k'}{\kappa}$  in the grid graph. There exist $z_i$ and $z_j$ such that $T(z_i) = \frac{k}{\kappa}$ and $T(z_j) = \frac{k'}{\kappa}$. By the triangle inequality,
	\begin{align*}
	\norm{z_i - z_j}_2 & \leq \norm{T(z_i) - z_i}_2 + \norm{T(z_i) - T(z_j)}_2 + \norm{T(z_j) - z_j}_2 \\
	& \leq 2C\left(\frac{\log n}{n}\right)^{1/d} + \frac{1}{n^{1/d}}
	\end{align*}
	and so by our choice of $r$, $i \sim j$ in $G_r$.
\end{proof}

To show the lower bound of \eqref{eqn: spectral_similarity_rgraph_grid}, we will make use of a technique from spectral graph theory known as Poincare's inequality.

\paragraph{Poincare inequality}

Let $G$ and $\widetilde{G}$ be undirected, unweighted graphs over vertex set $V$, with edge sets $E_G$ and $E_{\widetilde{G}}$, respectively. Let $\widetilde{\mathcal{P}}$ be the space of all paths over $E_{\widetilde{G}}$; that is, $\mathcal{P}$ consists of $\widetilde{P} \in \widetilde{\mathcal{P}}$ with
\begin{equation*}
\widetilde{P} = (\widetilde{e}_1, \ldots, \widetilde{e}_m) \tag{$\widetilde{e_i} \in E_{\widetilde{G}}$}
\end{equation*}
for some integer $m \geq 1$.

\begin{lemma}[Poincare inequality]
	\label{lem: poincare}
	Define a mapping $\gamma: E_G \to \mathcal{P}$ where for each $e = (\ell,\ell')$ in $E_G$
	\begin{equation*}
	\gamma(e) = ((\ell,u), \ldots, (v,\ell'))
	\end{equation*}
	meaning $e$ is mapped to a path which begins at $\ell$ and ends at $\ell'$. Then
	\begin{equation*}
	G \preceq \widetilde{G} \cdot \max_{e \in E_G} \abs{\gamma(e)}  \cdot b_{\gamma}
	\end{equation*}
	where $b_{\gamma}$ is a bottleneck parameter given by
	\begin{equation*}
	b_{\gamma} = \max_{\widetilde{e} \in E_{\widetilde{G}}} \abs{\set{e \in E: \widetilde{e} \in \gamma(e)}}
	\end{equation*}
\end{lemma}

Lemma \ref{lem: mass_transport_mapping} will allow us to construct such a mapping $\gamma$ from $E_r$ to $E_{\grid}$ and appropriately control parameters $\max_{e \in E_G} \abs{\gamma(e)}$ and $b_{\gamma}$.

\begin{lemma}
	There exists a mapping $\gamma: E_r \to \mathcal{P}_{\grid}$, the set of paths over $G_{\grid}$, such that the following quantities are bounded:
	\begin{enumerate}[(i)]
		\item \textit{Maximum path length.} 
		\begin{equation*}
		\max_{e \in E_G} \abs{\gamma(e)} \leq n^{1/d} \sqrt{d} \left(2C\left(\frac{\log n}{n}\right)^{1/d} + r \right)
		\end{equation*}
		\item \textit{Bottleneck.}
		\begin{equation*}
		b_{\gamma} \leq \left(n^{1/d} \sqrt{d} \left(2C\left(\frac{\log n}{n}\right)^{1/d} + r \right) \right)^{2d}
		\end{equation*}
	\end{enumerate}
	with probability at least $1 - n^{-\alpha}$ where $\alpha = c_1 (\log n)^{1/2}$ and $C, c_1 > 0$ are universal constants.
\end{lemma}

\begin{proof}	
	Assume $i \sim j$ in the graph $G_r$. By a similar set of steps to the above, we have
	\begin{equation*}
	\norm{T(z_i) - T(z_j)}_2 \leq 2C\left(\frac{\log t}{t}\right)^{1/d} + r
	\end{equation*}
	As a result, using the simple relation $\norm{x}_1 \leq \sqrt{d} \norm{x}_2$ for any $x \in \Rd$, we have
	\begin{equation*}
	\norm{T(z_i) - T(z_j)}_1 \leq \sqrt{d}(2C\left(\frac{\log t}{t}\right)^{1/d} + r)
	\end{equation*}
	Since each edge in the grid graph is of length $n^{1/d}$, it is easy to see that there exists a path between $T(z_i)$ and $T(z_j)$ in $G_{grid}$, $P(T(Z_i) \to T(Z_j))$ with no more than
	\begin{equation*}
	\frac{\sqrt{d}(2C\left(\frac{\log t}{t}\right)^{1/d} + r)}{t^{1/d}}
	\end{equation*}
	edges. The bound follows by Lemma \ref{lem: graph_ordering}.
\end{proof}

\section{Additional Theory and Proofs}

\subsection{Proof of Lemma \ref{lem: poincare}}

@@

\paragraph{Proof of Lemma \ref{lem: poincare}}
Let $G_e = (V, \set{e})$ and $P_e = (V, \set{\widetilde{e}: \widetilde{e} \in \gamma(e)})$ be the graphs associated with $e$ and $\gamma(e)$, respectively. By Lemma \ref{lem: path_poincare}, we have
\begin{equation*}
G_{e} \preceq \abs{P_e} P_e
\end{equation*}
Summing over all $e \in E_G$, we obtain
\begin{align*}
G & \preceq \sum_{e \in E_G} \abs{P_e} P_e \\
& \preceq \max_{e \in E_G} \abs{\gamma(e)} \sum_{e \in E_G} P_e \\
& \preceq \max_{e \in E_G} \abs{\gamma(e)} b_{\gamma}\cdot \widetilde{G}
\end{align*}
